### 2023-03-05 chapters/appendix/B-plan-of-work.tex

\thispagestyle{empty}
\chapter{Plan of work}
\pagenumbering{arabic}
\renewcommand*{\thepage}{B-\arabic{page}}

\section{Winter semester}

\begin{table}[h!]
\def\arraystretch{1.25}
\begin{tabular}{|l|p{12cm}|}
\hline
\textbf{Period} & \textbf{Work}                                                                                                                                                                                                                         \\ \hline
\nth{1} week         & Consultation with the supervisor on directions of the future work based on literature review during previous semester. Meeting schedule is established.
\\ \hline
\nth{2} week         & Outline of key sections of the analysis part of the thesis.
\\ \hline
\nth{3} week         & Match supporting literature with analysis sections. Further invesigation on the feature engineering in condition monitoring.
 \\ \hline
\nth{4} week         & TBD (Descibe machinery physics and maintanance strategies.)
 \\ \hline

\end{tabular}
\end{table}

\clearpage
\newpage

\section{Summer semester}

\clearpage