### 2023-03-30 chapters/02-analysis.tex
Extract sample frequency features and notes to separate file

\chapter{Problem analysis}

\section{Condition monitoring}

\subsection{Maintenance strategies}

\subsection{Vibration fault types}

\subsection{Technical standards}

\subsection{Sensor placement}

\section{Feature engineering}

\subsection{Feature extraction}

\subsection{Power spectral density segmentation}

\subsection{Feature selection}

\section{Diagnostics techniques}

\subsection{Novelty detection}

\section{IoT in Industry 4.0}