### 2023-04-12 chapters/02-analysis.tex
Compare spectra of different faults in MAFAULDA

\chapter{Problem analysis}

\section{Condition monitoring}
Why monitor with vibrations, Wear process curve, Fault - failure, Levels are dependant on load and lifecycle stage

\subsection{Maintenance strategies}
\paragraph{Reactive}
\paragraph{Preventive}
\paragraph{Predictive}
- value for the factory and assess risks asociated with potencial fault to assign them importance

\paragraph{Diagnosis indicators} - what we montitor? - when it fails and early signs of faulty component
- RUL (Remaining useful life) models - disadvantage lot of similiar machines (homogenous) or lots of runns until failure
\begin{itemize}
\item Similarity
\item Degradation - used by standards
\item Survival
\end{itemize}

\subsection{Vibration fault types}
- frequency ranges 1 - 300 Hz (shaft), 300 - 1000 Hz, 1000 - 10000 Hz (early bearings)
- Base analytical models: Jeffcott rotor - rotor dynamics, Bearnings model
- Resonance frequencies of each part - machine must run at speeds not aligned with resonance frequencies - Campbell diagram - task for mechanical engineers
- Faults - reasons and frequency content
\begin{itemize}
\item Synchrounous response - based on RPM
\item Mass unbalance
\item Misalignment
\item Eccentricity
\item Bent or bow shaft
\item Cracked shaft
\item Rotor rubs - friction
\item Looseness
\item Auxiliery mechanical systems: Gearbox, Bearings, Belt 
\end{itemize}

\subsection{Technical standards}
\paragraph{ISO 20816}
Part 1
\begin{itemize}
\item Measurement units - displacement, velocity, acceleration
\item RMS, and max. amplitude = severity
\item Measurement points for sensors (axial, radial) - image, and 45 degrees
\item Evaluation zones - A, B, C, D - Severity chart (Annex B) - Degradation model
\item Opeartional limits - Alarm, Trips
\end{itemize}

\paragraph{ISO 13373}
\begin{itemize}
\item Sensor mount type in relation to sensor resonance
\item Data presentation - standard display formats for analysis - trends, watefall plots ...
\item Potencial causes for faults (p. 45) - use in vibration fault types
\end{itemize}

\section{Feature engineering}
Large domain knowledge with compared to other areas of machine learning (mechanics - physics)
\subsection{Preprocessing}
\begin{itemize}
\item Detrending - DC removal filter
\item Time synchronous averaging
\end{itemize}

\subsection{Feature extraction}

\paragraph{Statistical measures}
\begin{itemize}
\item Standard Deviation
\item Max. amplitude
\item RMS amplitude
\item Skewness
\item Kurtosis \\
---
\item Spectral centroid
\item RMS frequency
\item Root variance frequency
\item Spectral kurtosis / Fast kurtogram
\item Harmonics (peaks)
\item Spectral Envelope
\item Harmonic spectral deviation \\
---
\item Energy
\item Spectral negentropy
\item TKEO - Teager-Kaiser energy operator
\end{itemize}

\paragraph{Signal decompositions - sparse approximations}
Matching pursuit algorithm optimalization problem
\begin{itemize}
\item FFT - Short Time Fourier Transform with Hamming window and Welch averaging (cite BP)
\item CWT-SST - Synchrosqueezing Wavelet Transform (vs. Transient-extracting transform)
\item WPD - Wavelet Packet Decomposition  - to approximation and detail coef. (Fejer-Korovkin wavelet)
\item EWT - Empirical Wavelet Transform - (Meyer wavelet) 
\end{itemize}

\subsection{Feature transformation}
\begin{itemize}
\item Principal Component Analysis (PCA)
\item Log transformation (Box-Cox Transform) to normal distribution
\item Normalization (min-max, standardize) 
\end{itemize}

\subsection{Feature selection}
Filter method - SelectKBest  in evaluation phase
\begin{itemize}
\item Variance Threshold
\item Pearson correlation
\item ANOVA F-value
\item Mutal information
\item Fisher score
\item Spectral feature selection algorithm (SPEC)
\end{itemize}

\section{Diagnostics techniques}
Idenification of faulty states in data streams in semi-supervised learning
\subsection{Novelty detection}
\begin{itemize}
\item Local Outlier Factor, Local Correlation Integral (Anomaly score)
\item DenStream (Density based clustering - DBSCAN)
\item Half-space Trees (Isolation forest)
\end{itemize}

\subsection{Classification}
\begin{itemize}
\item kNN + Metric Tree (M-Tree for neighbourhood queries) + Mahalanobis distance / Cosine similarity /  RBF similarity
\end{itemize}

\section{Evaluation Datasets}
(Pictures of machines)

\paragraph{MAFAULDA}\\
SpectraQuest's Machinery Fault Simulator (MFS) Alignment-Balance-Vibration (ABVT)
50 kHz, 5 sec. recordings, Imbalance, Horizontal/Vertical misalignment, Bearings (Overhang / Underhang) - Inner, Outer, Cage

\paragraph{CWRU}
2 HP (1.492 kW) Reliance Electric motor
Bearings - Inner, Outer
12 kHz, 48 kHz
fan and drive end bearings
Fault diameters of 7 mils, 14 mils, 21 mils, 28 mils, and 40 mils (1 mil=0.001 inches) in diameter were introduced separately at the inner raceway, rolling element (i.e. ball) and outer raceway. 
Faulted bearings were reinstalled into the test motor and vibration data was recorded for motor loads of 0 to 3 horsepower (motor speeds of 1797 to 1720 RPM).

\paragraph{Rotating Shaft}
Shaft -  unbalances of different sizes
4 kHz

\section{Sensor and microcontroller}