### 2023-04-28 thesis/chapters/02-analysis-B-features.tex

\section{Feature engineering}
Large domain knowledge with compared to other areas of machine learning (mechanics - physics)
\subsection{Preprocessing}
\begin{itemize}
\item Detrending - DC removal filter
\item Adaptive noise canellation of the background noise
\item Time synchronous averaging
\end{itemize}

\subsection{Feature extraction}
% Feature Engineering for Machine Learning
\cite{zheng_feature_2018}
% Feature Engineering and Selection: A Practical Approach for Predictive Models
\cite{johnson_feature_2019}
% A New Statistical Features Based Approach for Bearing Fault Diagnosis Using Vibration Signals
\cite{altaf_new_2022}
% A Novel Online Machine Learning Approach for Real-Time Condition Monitoring of Rotating Machines
\cite{mostafavi_novel_2021}
% Fault Detection of Bearing: An Unsupervised Machine Learning Approach Exploiting Feature Extraction and Dimensionality Reduction
\cite{brito_fault_2021}
% A large set of audio features for sound description
\cite{peeters_large_2004}
% Research on online intelligent monitoring system of band saw blade wear status based on multi‑feature fusion of acoustic emission signals
\cite{zhuo_research_2022}  
% Early Detection of Imbalance in Load and Machine In Front Load Washing Machines by Monitoring Drum Movement
\cite{mohammadi_early_2020}
% A Data Mining based Approach for Electric Motor Anomaly Detection Applied on Vibration Data
\cite{egaji_data_2020}
% Condition Monitoring with Vibration Signals
\cite{nandi_condition_2019}

% Vibration Analysis for IoT Enabled Predictive Maintenance
\cite{jung_vibration_2017}

\paragraph{Statistical measures}
\begin{itemize}
\item Standard Deviation
\item Max. amplitude
\item RMS amplitude
\item Skewness
\item Kurtosis \\
---
\item Spectral centroid
\item RMS frequency
\item Root variance frequency
\item Spectral kurtosis / Fast kurtogram
\item Harmonics (peaks) 
	% Comparative Study of Various CFAR Algorithms for Non-Homogenous Environments
	\cite{hatem_comparative_2018}
	% Multi-Scale Peak and Trough Detection Optimised for Periodic and Quasi-Periodic Neuroscience Data
	\cite{bishop_multi-scale_2018}
	% Non-Parametric Local Maxima and Minima Finder with Filtering Techniques for Bioprocess
	\cite{adikaram_non-parametric_2016}
	% Identification of harmonics and sidebands in a finite set of spectral components
	\cite{gerber_identification_2013}
	% Evaluation of Threshold-Based Algorithms for Detection of Spectral Peaks in Audio
	\cite{nunes_evaluation_2007}
	% Statistical techniques to select detection thresholds for peak signals in ice-core data
	\cite{karlof_statistical_2005}
\item Spectral Envelope
\item Harmonic spectral deviation \\
---
\item Energy
\item Spectral negentropy
	% Spectral negentropy and kurtogram performance comparison for bearing fault diagnosis
	\cite{avoci_spectral_2020}
\item TKEO - Teager-Kaiser energy operator
	%Application of Teager–Kaiser Energy Operator in the Early Fault Diagnosis of Rolling Bearings
	\cite{shi_application_2022}
\end{itemize}

\paragraph{Signal decompositions - sparse approximations}
Matching pursuit algorithm optimalization problem
\cite{song_mfbd_2021}
\begin{itemize}
\item FFT - Short Time Fourier Transform with Hamming window and Welch averaging
	% \cite{oulmane_automatic_2015}

\item CWT-SST - Synchrosqueezing Wavelet Transform (vs. Transient-extracting transform) 
	% The fast continuous wavelet transformation (fCWT) for real-time, high-quality, noise-resistant time–frequency analysis
	\cite{arts_fast_2022}
	% A Concentrated Time–Frequency Analysis Tool for Bearing Fault Diagnosis
	\cite{yu_concentrated_2020}
	% Applications of the synchrosqueezing transform in seismic time-frequency analysis
	\cite{herrera_applications_2014}

\item WPD - Wavelet Packet Decomposition  - to approximation and detail coef. (Fejer-Korovkin wavelet)  	
	% Wavelet Packet Feature Extraction for Vibration Monitoring		
	\cite{yen_wavelet_2000}
	% A wavelet approach to dimension reduction and classiﬁcation of hyperspectral data
	\cite{wickmann_wavelet_2007}

\item EWT - Empirical Wavelet Transform - (Meyer wavelet)
	% On the computational complexity of the empirical mode decomposition algorithm
	\cite{wang_computational_2014}
	% Novel self-adaptive vibration signal analysis: Concealed component decomposition and its application in bearing fault diagnosis
	\cite{tiwari_novel_2021}
	% The MFBD: a novel weak features extraction method for rotating machinery
	\cite{song_mfbd_2021}
	% Fault Feature Extraction and Enhancement of Rolling Element Bearings Based on Maximum Correlated Kurtosis Deconvolution and Improved Empirical Wavelet Transform
	\cite{li_fault_2019}
	% An Improved Empirical Wavelet Transform for Noisy and Non-Stationary Signal Processing
	\cite{zhuang_improved_2020}
	% Time and frequency domain scanning fault diagnosis method based on spectral negentropy and its application
	\cite{yonggang_time_2020}
	% An Adaptive Spectrum Segmentation Method to Optimize Empirical Wavelet Transform for Rolling Bearings Fault Diagnosis
	\cite{xu_adaptive_2019}
	% Improved empirical wavelet transform (EWT) and its application in non‑stationary vibration signal of transformer
	\cite{ni_improved_2022}
\end{itemize}

\subsection{Feature transformation}
\cite{zheng_feature_2018}
\begin{itemize}
\item Principal Component Analysis (PCA)
\item Log transformation (Box-Cox Transform) to normal distribution
\item Normalization (min-max, standardize) 
\end{itemize}

\subsection{Feature selection}
\cite{nandi_condition_2019}
Filter method - SelectKBest  in evaluation phase
\begin{itemize}
\item Variance Threshold
\item Pearson correlation
\item ANOVA F-value
\item Mutal information
\item Fisher score
\item Spectral feature selection algorithm (SPEC)
\end{itemize}