### 2023-05-07 thesis/chapters/02-analysis-C-diagnostics.tex
Summarize articles of high relevance for feature engineering section

\section{Diagnostics techniques}
Idenification of faulty states in data streams in semi-supervised learning

% Analysis of different RNN autoencoder variants for time series classification and machine prognostics
\cite{yu_analysis_2021}

\subsection{Novelty detection}
\cite{gervasi_anomaly_2020}

\begin{itemize}
\item Local Outlier Factor, Local Correlation Integral (Anomaly score)
	% One-Class Classification with LOF and LOCI: An Empirical Comparison
	\cite{janssens_one-class_2007}
	% Designing a Streaming Algorithm for Outlier Detection in Data Mining - An Incremental Approach
	\cite{yu_designing_2020}

\item DenStream (Density based clustering - DBSCAN)
	% Density-Based Clustering over an Evolving Data Stream with Noise
	\cite{cao_density-based_2006}
	% A Modified Approach of OPTICS Algorithm for Data Streams
	\cite{shukla_modified_2017}
	% Data Clustering - Algorithms and Applications
	\cite{aggarwal_data_2014}
	% State-of-the-art on clustering data streams
	\cite{ghesmoune_state---art_2016}
	% Cluster-Reduce: Compressing Sketches for Distributed Data Streams
	\cite{zhao_cluster-reduce_2021}

\item Half-space Trees (Isolation forest)
	% Fast Anomaly Detection for Streaming Data
	\cite{tan_fast_2011}
	% Anomaly Detection for Data Streams Based on Isolation Forest using Scikit-multiflow
	\cite{gervasi_anomaly_2020}

\end{itemize}

\subsection{Classification}
\begin{itemize}
\item kNN + Metric Tree (M-Tree for neighbourhood queries) + Euclidian Mahalanobis distance /  RBF similarity

% Review of Artificial Intelligence-based Bearing Vibration Monitoring
\cite{sheng_review_2020}
% Semi-Supervised Learning on Data Streams via Temporal Label Propagation
\cite{wagner_semi-supervised_2018}
% Minimum covariance determinant and extensions
\cite{hubert_minimum_2018}
\end{itemize}

% Feature-based performance of SVM and KNN classifiers for diagnosis of rolling element bearing faults
\cite{jamil_feature-based_2021}
% Classification of washing machines vibration signals using discrete wavelet analysis for feature extraction
\cite{goumas_classification_2002}
% Semi-Supervised Learning
\cite{chapelle_semi-supervised_2006}

% A Novel Online Machine Learning Approach for Real-Time Condition Monitoring of Rotating Machines
\cite{maurya_condition-based_2021}