### 2023-10-31 thesis/chapters/appendix/B-plan-of-work.tex
Revise plan of work

\thispagestyle{empty}
\chapter{Plan of work}
\pagenumbering{arabic}
\renewcommand*{\thepage}{B-\arabic{page}}

\section{Summer semester - DP1}

\begin{table}[h!]
\def\arraystretch{1.25}
\begin{tabular}{|l|p{12cm}|}
\hline
\textbf{Period} & \textbf{Work}                                                                                                                                                                                                                         \\ \hline
\nth{1} week         & Consultation with the supervisor on directions of the future work based on literature review during previous semester.
\\ \hline
\nth{2} week         & Outline the key sections of the analysis part in the thesis.
\\ \hline
\nth{3} week         & Match supporting literature with analysis sections. Further invesigation on the feature engineering methodology in CbM.
 \\ \hline
\nth{4} week         & Summarize notes from condition monitoring articles and videorecordings of tutorials and conferences.
 \\ \hline
\nth{5} week         & Research transformation of vibration signal to feature space using time-frequency, harmonic and energy statistical metrics. Progress report meeting with the supervisor.
 \\ \hline
\nth{6} week         & Find articles and take notes about unsupervised and semi-supervised techniques in streaming data for machinery diagnostics.
 \\ \hline
\nth{7} week         & Narrow down wide variety applicable methods for signal decomposition.
 \\ \hline
 \nth{8} week         & Exploratory analysis on evaluation datasets. Progress report meeting with the supervisor to topic of related work.
 \\ \hline
 \nth{9} week         & Organize detailed outline out of notes gathered during literature research. 
 \\ \hline
  \nth{10} week         & Write up the problem analysis about condition monitoring and evaluation datasets.
 \\ \hline
  \nth{11} week         & Write up the analysis section about feature engineering.
 \\ \hline
  \nth{12} week         & Write up the analysis section about machine learning diagnostics and consult final choice of methods in the analysis section.
 \\ \hline
\end{tabular}
\end{table}

\clearpage
\newpage

\section{Winter semester - DP2}

\begin{table}[h!]
\def\arraystretch{1.25}
\begin{tabular}{|l|p{12cm}|}
\hline
\textbf{Period} & \textbf{Work}                                                                                                                                                                                                                         \\ \hline
\nth{1} week         & Semester kickoff meeting with the supervisor and consultant to set goals, experiments and discuss status of collaborations with the industry.
\\ \hline
\nth{2} week         &  Arrange collaboration with alternative industry partner. Prepare checklist for technical inspection of machinery. Construct device for exploratory measurements.
\\ \hline
\nth{3} week         & Technical inspection of air conditioning units in data centre of the industry partner. Feature extraction step to calculate features from Mafaulda dataset.
 \\ \hline
\nth{4} week         & Consultation with the supervisor about choice of machine to measure (scroll compressor), device for measurement, and requirements for dataset gathering. Fine tune measurement device.
 \\ \hline
\nth{5} week         &  In feature selection use various metrics to determine sets of best data features in Mafaulda dataset.
 \\ \hline
\nth{6} week         & Generate reports from jupyter notebooks. Feature sampling used in kNN multiclass and binary classificator.
 \\ \hline
\nth{7} week         & Explore streaming ML models with Mafaulda dataset. Consultation with the supervisor and status report.
 \\ \hline
 \nth{8} week         & ?Anomaly detection models evaluation and conclude/gather all the findings.
 \\ \hline
 \nth{9} week         & ??Make further exploratory measurements with rewritten code ensuring more precise accelerometer sampling. Make a plan of measurement and negotiate it with industry partner.
 \\ \hline
  \nth{10} week         & ?Begin writing implementation
 \\ \hline
  \nth{11} week         & 
 \\ \hline
  \nth{12} week         & 
 \\ \hline
\end{tabular}
\end{table}

\clearpage
\newpage

\begin{table}[h!]
\def\arraystretch{1.25}
\begin{tabular}{|l|p{12cm}|}
\hline
\textbf{Period} & \textbf{Work}                                                                                                                                                                                                                         \\ \hline
\nth{1}  - \nth{4} week         & Apply feature engineering and classifications methods to evalution datasets
\\ \hline
\nth{4}  - \nth{8} week         & Design the vibration measurement for factory according to technical standards. 
\\ \hline
\nth{8} - \nth{12} week         & Take preliminary measurements in the factory and compare signals with evaluation datasets.
 \\ \hline
\end{tabular}
\end{table}

\section{Summer semester - DP3}
\begin{table}[h!]
\def\arraystretch{1.25}
\begin{tabular}{|l|p{12cm}|}
\hline
\textbf{Period} & \textbf{Work}                                                                                                                                                                                                                         \\ \hline
\nth{1}  - \nth{4} week         & 
\\ \hline
\nth{4}  - \nth{8} week         & 
\\ \hline
\nth{8} - \nth{12} week         & 
 \\ \hline
\end{tabular}
\end{table}

\clearpage