### 2023-12-20 thesis/chapters/preliminaries/annotation.tex
Principal components analysis

\thispagestyle{empty}
\section*{Annotation}
\UniversityEN \\
\uppercase{\FacultyEN}
\vspace{-8pt}
{\setlength{\mathindent}{0cm}
\begin{align*}
&\text{Degree course:} && \text{\StudyProgrammeEN} \\
&\text{Author:} && \text{\AuthorEN} \\
&\text{\ThesisEN:} && \text{\TitleEN} \\
&\text{Supervisor:} && \text{\SupervisorEN} \\
&\text{Departmental advisor:} && \text{\DepartmentalAdvisorEN} \\
&\text{Consultant:} && \text{\ConsultantEN} \\
&\text{\DateEN}
\end{align*}}

\emptypage

\thispagestyle{empty}
\section*{Anotácia}
\University \\
\uppercase{\Faculty}
\vspace{-8pt}
{\setlength{\mathindent}{0cm}
\begin{align*}
&\text{Študijný program:} && \text{\StudyProgramme} \\
&\text{Autor:} && \text{\Author} \\
&\text{\Thesis:} && \text{\Title} \\
&\text{Vedúci diplomovej práce:} && \text{\Supervisor} \\
&\text{Pedagogický vedúci:} && \text{\DepartmentalAdvisor} \\
&\text{Konzultant:} && \text{\Consultant} \\
&\text{\Date}
\end{align*}}

\emptypage