### 2023-12-26 thesis/chapters/appendix/C-measurement-plan.tex
Vibration measurement plan

\thispagestyle{empty}
\chapter{Data collection plan} \label{appendix:data-collection}
\pagenumbering{arabic}
\renewcommand*{\thepage}{C-\arabic{page}}

The vibration measurements will occur biweekly and will contain three trials for each position as designed in the checklist spreadsheet below. The sensor is mounted to the machine with adhesive on double-sided carpet tape. After every trial, the sensor is detached and attached again.

Each triaxial recording has a 60-second duration at a sampling frequency of 26.6 kHz and a sample resolution of 16 bits. Measurement places are described according to the MIMOSA convention (ISO~13373-1). The total space requirement for a complete dataset composed out of 192 recordings is 1.71 GiB.

\begin{table}[ht]
\centering
\renewcommand{\arraystretch}{1.2}
\begin{tabular}{|l|l|lll|lll|lll|lll|}
\hline
\textbf{Machine}                                                                       & \textbf{Placement / Inspection} & \multicolumn{3}{l|}{\textbf{\nth{1}}}                & \multicolumn{3}{l|}{\textbf{\nth{2}}}                & \multicolumn{3}{l|}{\textbf{\nth{3}}}                & \multicolumn{3}{l|}{\textbf{\nth{4}}}                \\ \hline
\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Compressor\\ AC no.3\end{tabular}}} & SFTA001AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & SFTA002AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \hline
\multirow{2}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Compressor\\ AC no.5\end{tabular}}} & SFTA001AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & SFTA002AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \hline
\multirow{4}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Pump\\ KSB (new)\end{tabular}}}     & MTRA001AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & MTRA002AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & PMPA003AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & PMPA004AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \hline
\multirow{4}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Pump\\ Sigma (big)\end{tabular}}}   & MTRA001AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & MTRA002AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & PMPA003AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & PMPA004AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \hline
\multirow{4}{*}{\textbf{\begin{tabular}[c]{@{}l@{}}Pump\\ Sigma (small)\end{tabular}}} & MTRA001AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & MTRA002AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & PMPA003AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \cline{2-14} 
                                                                                       & PMPA004AT000TN                               & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} &  \\ \hline
\end{tabular}
\end{table}

\clearpage