### 2024-04-13 thesis/chapters/02-analysis.tex
Refactor jupyter

\chapter{Problem analysis} \label{chapter:problem-analysis}
In the problem analysis chapter we explore the feature engineering methods and machine learning algorithms for fault diagnostics. The basis we build upon is the domain knowledge of mechanical engineers in vibration signal measurement and its evaluation.

\input{chapters/analysis/1-condition-monitoring}
\input{chapters/analysis/2-features}
\input{chapters/analysis/3-diagnostics}
\input{chapters/analysis/4-datasets}