### 2024-04-13 thesis/chapters/appendix/C-user-guide.tex

\thispagestyle{empty}
\chapter{User Guide} \label{appendix:user-guide}
\pagenumbering{arabic}
\renewcommand*{\thepage}{C-\arabic{page}}