### 2024-04-13 thesis/chapters/preliminaries/acknowledgments.tex

% Poďakovanie
\thispagestyle{empty}
\vspace*{\fill}

\begin{center}
\settowidth\longest{\itshape ---Wir müssen wissen.---}
\parbox{\longest}{
  \hrulefill\hspace{0.2cm} \decofourleft\decofourright \hspace{0.2cm} \hrulefill\par
  \raggedright{
  \itshape
  	Wir müssen wissen. \\ Wir werden wissen.\par
  }   
  \raggedleft{--- David Hilbert}\par
  \hrulefill\hspace{0.2cm} \decofourleft\decofourright\hspace{0.2cm} \hrulefill\par
}
\end{center}

\vspace*{\fill}
\section*{Poďakovanie}
{\small TODO }
%Ing. Marcel Baláž, PhD.
%R-DAS, s.r.o.
%Ing. Lukáš Doubravský
%
%VNET, a.s.
%Michal Ország
%Mgr. Vladimír Kupčo
%
%BVS, a.s.
%Ing. Peter Csóka
%Ing. Peter Kmeťko
%
%SjF STU:
%prof. Ing. Stanislav Žiaran, CSc.
%Ing. Ondrej Chlebo, PhD.

% rodina
% kamaráti-spolužiaci-kolegovia, hlavne:
% Tomáš Starovič
% Mgr. Ivona Hrivová
% Bc. Sofia Mlynarčíková
% Bc. Ladislav Gaža
% Bc. Michaela Gubovská
% Bc. Jakub Hajdu
% Ing. Michal Juranyi

\vspace{3cm}