### 2024-04-29 thesis/chapters/05-evaluation.tex
Use bearing A instead of A+B

\chapter{Evaluation} \label{section:evaluation}

\section{MaFaulDa in k-nearest neighbors}

\subsection{Complete feature sets}
% Confusion matrix
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/TD-confusion-matrix.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/FD-confusion-matrix.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \caption{Confusion matrix for complete sets of features}
\end{figure}

% Nearest neighbors - how many points have neighbourhood of same class A+B
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/neighborhood.png}
        \caption{No severity}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/neighborhood-severity.png}
        \caption{Severity}
    \end{subfigure}
    \caption{Neighborhood with same label}
\end{figure}

% Accuracy based on scenario
% No optimal K - data is noisy
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/TD-severity.png}
        \caption{Time-domain features (severity)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/all-features/FD-severity.png}
        \caption{Frequency-domain features (severity)}
    \end{subfigure} 
    \caption{Accuracy dependency on k-value}
\end{figure}

\subsection{Feature subset combinations}
% Boxplots
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/TD-3-A-False-False-F3.png}
        \caption{Time domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/TD-3-A-False-False-K5.png}
        \caption{Time domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/FD-3-A-False-False-F3.png}
        \caption{Frequency domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/FD-3-A-False-False-K5.png}
        \caption{Frequency domain features}
    \end{subfigure}
    \caption{Model accuracy distribution (Bearings A+B, Dim = 3, Severity False)}
\end{figure}

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/TD-3-A-True-False-F3.png}
        \caption{Time domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/TD-3-A-True-False-K5.png}
        \caption{Time domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/FD-3-A-True-False-F3.png}
        \caption{Frequency domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/FD-3-A-True-False-K5.png}
        \caption{Frequency domain features}
    \end{subfigure}
    \caption{Model accuracy distribution (Severity True)}
\end{figure}

\subsection{Feature selection techniques}
% Accuracy tables
% Compare linear complexity accuracy with whole distribution

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-TD-train.png}
        \caption{Time domain features (train)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-TD-test.png}
        \caption{Time domain features (test)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-FD-train.png}
        \caption{Frequency domain features (train)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/model-distr-fsel-k5-f3-FD-test.png}
        \caption{Frequency domain features (test)}
    \end{subfigure}
    \caption{Model distibution with feature selection methods (k = 5, f = 3)}
\end{figure}

% compare dimensions (table)
\begin{table}[h]
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{|l|rr|rr|r|l|}
\hline
\multirow{2}{*}{\textbf{Feature set}} & \multicolumn{2}{l|}{\textbf{Accuracy}}                                   & \multicolumn{2}{l|}{\textbf{Percentile}}                                 & \multicolumn{1}{l|}{\multirow{2}{*}{\textbf{Axis}}} & \multirow{2}{*}{\textbf{Best features}} \\ \cline{2-5}
                                      & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{}                               &                                         \\ \hline
All features                          & \multicolumn{1}{r|}{92.52}          & 87.68                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 1                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{87.90}          & 79.42                              & \multicolumn{1}{r|}{96.67}          & 96.67                              & 1                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{89.70}          & 81.96                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 1                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{87.51}          & 79.35                              & \multicolumn{1}{r|}{95.83}          & 96.67                              & 1                                                   & zerocross, shape, pp                    \\ \hline
Correlation                           & \multicolumn{1}{r|}{87.51}          & 79.35                              & \multicolumn{1}{r|}{95.83}          & 96.67                              & 1                                                   & shape, zerocross, pp                    \\ \hline
F statistic                           & \multicolumn{1}{r|}{86.95}          & 78.65                              & \multicolumn{1}{r|}{88.33}          & 95.00                              & 1                                                   & pp, zerocross, aac                      \\ \hline
Mutual information                    & \multicolumn{1}{r|}{86.98}          & 78.28                              & \multicolumn{1}{r|}{89.17}          & 94.17                              & 1                                                   & zerocross, shape, rms                   \\ \hline
All features                          & \multicolumn{1}{r|}{96.03}          & 92.80                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 3                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{91.20}          & 84.67                              & \multicolumn{1}{r|}{95.00}          & 93.33                              & 3                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{91.93}          & 85.47                              & \multicolumn{1}{r|}{100.00}         & 99.17                              & 3                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{90.59}          & 84.07                              & \multicolumn{1}{r|}{91.67}          & 90.00                              & 3                                                   & zerocross, pp, rms                      \\ \hline
Correlation                           & \multicolumn{1}{r|}{91.63}          & 85.04                              & \multicolumn{1}{r|}{99.17}          & 95.83                              & 3                                                   & zerocross, shape, pp                    \\ \hline
F statistic                           & \multicolumn{1}{r|}{90.59}          & 84.07                              & \multicolumn{1}{r|}{91.67}          & 90.00                              & 3                                                   & rms, pp, zerocross                      \\ \hline
Mutual information                    & \multicolumn{1}{r|}{91.63}          & 85.04                              & \multicolumn{1}{r|}{99.17}          & 95.83                              & 3                                                   & zerocross, pp, shape                    \\ \hline
\end{tabular}
\end{adjustbox}
\caption{k=5, f=3, bearing=A, severity=False, domain=TD}
\end{table}

\begin{table}[h]
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{|l|rr|rr|r|l|}
\hline
\multirow{2}{*}{\textbf{Feature set}} & \multicolumn{2}{l|}{\textbf{Accuracy}}                                   & \multicolumn{2}{l|}{\textbf{Percentile}}                                 & \multicolumn{1}{l|}{\multirow{2}{*}{\textbf{Axis}}} & \multirow{2}{*}{\textbf{Best features}} \\ \cline{2-5}
                                      & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{}                               &                                         \\ \hline
All features                          & \multicolumn{1}{r|}{88.15}          & 79.99                              & \multicolumn{1}{r|}{99.39}          & 98.18                              & 1                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{83.83}          & 74.43                              & \multicolumn{1}{r|}{77.58}          & 83.03                              & 1                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{88.93}          & 81.86                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 1                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{86.87}          & 77.51                              & \multicolumn{1}{r|}{95.76}          & 93.33                              & 1                                                   & entropy, roll\_off, flux                \\ \hline
Correlation                           & \multicolumn{1}{r|}{80.54}          & 68.01                              & \multicolumn{1}{r|}{33.33}          & 31.52                              & 1                                                   & kurtosis, roll\_off, skewness           \\ \hline
F statistic                           & \multicolumn{1}{r|}{86.87}          & 77.51                              & \multicolumn{1}{r|}{95.76}          & 93.33                              & 1                                                   & flux, entropy, roll\_off                \\ \hline
Mutual information                    & \multicolumn{1}{r|}{88.11}          & 80.96                              & \multicolumn{1}{r|}{99.39}          & 99.39                              & 1                                                   & entropy, centroid, roll\_off            \\ \hline
All features                          & \multicolumn{1}{r|}{93.67}          & 88.45                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 3                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{86.76}          & 78.51                              & \multicolumn{1}{r|}{64.85}          & 70.91                              & 3                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{92.86}          & 87.52                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 3                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{90.70}          & 84.74                              & \multicolumn{1}{r|}{94.55}          & 96.36                              & 3                                                   & roll\_off, flux, centroid               \\ \hline
Correlation                           & \multicolumn{1}{r|}{84.55}          & 75.10                              & \multicolumn{1}{r|}{38.18}          & 42.42                              & 3                                                   & roll\_off, negentropy, skewness         \\ \hline
F statistic                           & \multicolumn{1}{r|}{90.70}          & 84.74                              & \multicolumn{1}{r|}{94.55}          & 96.36                              & 3                                                   & roll\_off, flux, centroid               \\ \hline
Mutual information                    & \multicolumn{1}{r|}{90.73}          & 83.60                              & \multicolumn{1}{r|}{94.55}          & 94.55                              & 3                                                   & roll\_off, entropy, noisiness           \\ \hline
\end{tabular}
\end{adjustbox}
\caption{k=5, f=3, bearing=A, severity=False, domain=FD}
\end{table}

\begin{table}[h]
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{|l|rr|rr|r|l|}
\hline
\multirow{2}{*}{\textbf{Feature set}} & \multicolumn{2}{l|}{\textbf{Accuracy}}                                   & \multicolumn{2}{l|}{\textbf{Percentile}}                                 & \multicolumn{1}{l|}{\multirow{2}{*}{\textbf{Axis}}} & \multirow{2}{*}{\textbf{Best features}} \\ \cline{2-5}
                                      & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{}                               &                                         \\ \hline
All features                          & \multicolumn{1}{r|}{94.99}          & 92.07                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 1                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{92.67}          & 89.20                              & \multicolumn{1}{r|}{84.17}          & 87.50                              & 1                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{93.99}          & 91.31                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 1                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{92.71}          & 88.55                              & \multicolumn{1}{r|}{85.00}          & 80.83                              & 1                                                   & zerocross, shape, rms                   \\ \hline
Correlation                           & \multicolumn{1}{r|}{92.71}          & 88.55                              & \multicolumn{1}{r|}{85.00}          & 80.83                              & 1                                                   & zerocross, shape, rms                   \\ \hline
F statistic                           & \multicolumn{1}{r|}{92.71}          & 88.55                              & \multicolumn{1}{r|}{85.00}          & 80.83                              & 1                                                   & zerocross, shape, rms                   \\ \hline
Mutual information                    & \multicolumn{1}{r|}{91.31}          & 86.99                              & \multicolumn{1}{r|}{67.50}          & 67.50                              & 1                                                   & zerocross, shape, impulse               \\ \hline
All features                          & \multicolumn{1}{r|}{96.42}          & 94.76                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 3                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{94.63}          & 92.07                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 3                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{94.58}          & 91.71                              & \multicolumn{1}{r|}{100.00}         & 99.17                              & 3                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{94.31}          & 91.40                              & \multicolumn{1}{r|}{98.33}          & 95.83                              & 3                                                   & zerocross, shape, rms                   \\ \hline
Correlation                           & \multicolumn{1}{r|}{94.31}          & 91.40                              & \multicolumn{1}{r|}{98.33}          & 95.83                              & 3                                                   & zerocross, shape, rms                   \\ \hline
F statistic                           & \multicolumn{1}{r|}{94.31}          & 91.40                              & \multicolumn{1}{r|}{98.33}          & 95.83                              & 3                                                   & shape, zerocross, rms                   \\ \hline
Mutual information                    & \multicolumn{1}{r|}{91.90}          & 88.51                              & \multicolumn{1}{r|}{72.50}          & 76.67                              & 3                                                   & zerocross, shape, clearance             \\ \hline
\end{tabular}
\end{adjustbox}
\caption{k=5, f=3, bearing=A, severity=True, domain=TD}
\end{table}

\begin{table}[h]
\begin{adjustbox}{width=\textwidth}
\begin{tabular}{|l|rr|rr|r|l|}
\hline
\multirow{2}{*}{\textbf{Feature set}} & \multicolumn{2}{l|}{\textbf{Accuracy}}                                   & \multicolumn{2}{l|}{\textbf{Percentile}}                                 & \multicolumn{1}{l|}{\multirow{2}{*}{\textbf{Axis}}} & \multirow{2}{*}{\textbf{Best features}} \\ \cline{2-5}
                                      & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{\textbf{Train}} & \multicolumn{1}{l|}{\textbf{Test}} & \multicolumn{1}{l|}{}                               &                                         \\ \hline
All features                          & \multicolumn{1}{r|}{94.17}          & 90.94                              & \multicolumn{1}{r|}{100.00}         & 99.39                              & 1                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{91.49}          & 87.78                              & \multicolumn{1}{r|}{80.00}          & 86.67                              & 1                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{94.13}          & 91.37                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 1                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{93.75}          & 90.29                              & \multicolumn{1}{r|}{99.39}          & 99.39                              & 1                                                   & entropy, centroid, flux                 \\ \hline
Correlation                           & \multicolumn{1}{r|}{93.75}          & 90.29                              & \multicolumn{1}{r|}{99.39}          & 99.39                              & 1                                                   & entropy, flux, centroid                 \\ \hline
F statistic                           & \multicolumn{1}{r|}{93.75}          & 90.29                              & \multicolumn{1}{r|}{99.39}          & 99.39                              & 1                                                   & entropy, centroid, flux                 \\ \hline
Mutual information                    & \multicolumn{1}{r|}{93.75}          & 90.29                              & \multicolumn{1}{r|}{99.39}          & 99.39                              & 1                                                   & centroid, entropy, flux                 \\ \hline
All features                          & \multicolumn{1}{r|}{95.20}          & 93.20                              & \multicolumn{1}{r|}{100.00}         & 100.00                             & 3                                                   &                                         \\ \hline
PCA PC                                & \multicolumn{1}{r|}{92.14}          & 88.84                              & \multicolumn{1}{r|}{69.09}          & 73.94                              & 3                                                   &                                         \\ \hline
Best features                         & \multicolumn{1}{r|}{94.64}          & 91.94                              & \multicolumn{1}{r|}{100.00}         & 99.39                              & 3                                                   &                                         \\ \hline
Rank product                          & \multicolumn{1}{r|}{94.50}          & 92.19                              & \multicolumn{1}{r|}{99.39}          & 100.00                             & 3                                                   & entropy, centroid, flux                 \\ \hline
Correlation                           & \multicolumn{1}{r|}{94.50}          & 92.19                              & \multicolumn{1}{r|}{99.39}          & 100.00                             & 3                                                   & entropy, centroid, flux                 \\ \hline
F statistic                           & \multicolumn{1}{r|}{94.50}          & 92.19                              & \multicolumn{1}{r|}{99.39}          & 100.00                             & 3                                                   & entropy, flux, centroid                 \\ \hline
Mutual information                    & \multicolumn{1}{r|}{93.32}          & 90.67                              & \multicolumn{1}{r|}{90.91}          & 91.52                              & 3                                                   & roll\_off, entropy, noisiness           \\ \hline
\end{tabular}
\end{adjustbox}
\caption{k=5, f=3, bearing=A, severity=True, domain=FD}
\end{table}

% compare percentiles
% Feature combinations accuracy
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/fsel-accuracy.png}
        \caption{Accuracy}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-combinations/fsel-percentile.png}
        \caption{Percentile}
    \end{subfigure}
    \caption{Feature selection method choices}
\end{figure}

% In fsel percentile - mention median from notebook for each method - table
\begin{table}[h]
\centering
\begin{tabular}{|l|r|r|}
\hline
\textbf{Method}    & \multicolumn{1}{l|}{\textbf{Median percentile {[}\%{]}}} & \multicolumn{1}{l|}{\textbf{Median accuracy {[}\%{]}}} \\ \hline
Rank product       & 93.94                                                    & 81.84                                                  \\ \hline
Mutual information & 93.33                                                    & 81.33                                                  \\ \hline
F statistic        & 91.01                                                    & 81.26                                                  \\ \hline
Correlation        & 84.11                                                    & 79.25                                                  \\ \hline
\end{tabular}
\end{table}

% Best feature selection methods table (order by best count)
\begin{table}[h]
\centering
\begin{tabular}{|l|r|r|r|}
\hline
                            & \multicolumn{1}{l|}{\textbf{Best in scenarios}} & \multicolumn{1}{l|}{\textbf{Scenarios {[}\%{]}}} & \multicolumn{1}{l|}{\textbf{Mean percentile {[}\%{]}}} \\ \hline
\textbf{Rank product}       & 151                                                     & 69.91                                            & 90.85                                                  \\ \hline
\textbf{Mutual information} & 46                                                      & 21.30                                            & 90.35                                                  \\ \hline
\textbf{Correlation}        & 13                                                      & 6.02                                             & 88.88                                                  \\ \hline
\textbf{F statistic}        & 6                                                       & 2.78                                             & 78.70                                                  \\ \hline
\textbf{$\Sigma$}           & 216                                                     & 100                                       & -                                                      \\ \hline
\end{tabular}
\end{table}

% 2d and 3d plots of best features
\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/labels/TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/labels/FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/labels/TD-severity.png}
        \caption{Time-domain features (severity)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/labels/FD-severity.png}
        \caption{Frequency-domain features (severity)}
    \end{subfigure} 
    \caption{3 Best features and their ranges}
\end{figure}

\subsection{Incremental learning}
% Ordering of severity
\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/order-natural.png}
        \caption{Order by raising severity}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/order-severity.png}
        \caption{Order after relabeing normal class}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/severity-levels.png}
        \caption{Order of severities}
    \end{subfigure}
    \caption{Ordering of faults}
\end{figure}

% Gradual
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/gradual-TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/gradual-FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/gradual-TD-severity.png}
        \caption{Time-domain features (severity)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/gradual-TD-severity.png}
        \caption{Frequency-domain features (severity)}
    \end{subfigure} 
    \caption{Gradual learning (window = 1, skip = 0)}
\end{figure}

% Tumbling windows
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/tumbling-TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/tumbling-FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/tumbling-TD-severity.png}
        \caption{Time-domain features (severity)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/tumbling-FD-severity.png}
        \caption{Frequency-domain features (severity)}
    \end{subfigure} 
    \caption{Tumbling window in incremental learning}
\end{figure}

% % Label skips and parts of dataset
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/all-features-TD-skip.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/all-features-FD-skip.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-labels-TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-labels-FD.png}
        \caption{Frequency-domain features}
    \end{subfigure} 
    \caption{Skip labels in incremental learning}
\end{figure}

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-TD.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-FD.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-TD-severity.png}
        \caption{Time-domain features (severity)}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/incremental-learning/skip-label-FD-severity.png}
        \caption{Frequency-domain features (severity)}
    \end{subfigure} 
    \caption{Skip labels in incremental learning with different sizes}
\end{figure}

\section{Validation of data logger}
% Validation of firmware on standing fan - speeds
\begin{figure}[ht]
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-1.png}
        \caption{Speed 1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-2.png}
        \caption{Speed 2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.3\textwidth}
        \includegraphics[width=\textwidth]{assets/results/standing-fan/fan-audio-speed-3.png}
        \caption{Speed 3}
    \end{subfigure}
    \caption{Frequency spectrum from sound of standing fan}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.5\textwidth]{assets/results/standing-fan/standing-fan-accel.png}
    \caption{Estimation of rotation speed from vibrations}
\end{figure}

\section{Industrial dataset analysis}

\subsection{Feature ranges}
% Amplitude histograms - 20m/s^2, 40m/s^2, 150m/s^
% Histograms of machinery measurements

% Box plots of features
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/pumps-TD-dim-3.png}
        \caption{Time-domain features}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{\textwidth}
        \includegraphics[width=\textwidth]{assets/results/feature-values/pumps-FD-dim-3.png}
        \caption{Frequency-domain features}
    \end{subfigure}
    \caption{Feature range in pumps}
\end{figure}

\subsection{Time-frequency waveforms}
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{assets/results/eda/frequency-spectrum-motors.png}
        \caption{Motors}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{assets/results/eda/frequency-spectrum-pumps.png}
        \caption{Pumps}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.32\textwidth}
        \includegraphics[width=\textwidth]{assets/results/eda/frequency-spectrum-compressors.png}
        \caption{Compressor}
    \end{subfigure} 
    \caption{Frequency domain}
\end{figure}

% Time-frequency spectrum

% Low frequencies (under 1KHz)

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/time-frequency-spectrum/K3-z-STFT-1kHz.png}
        \caption{K3}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/time-frequency-spectrum/M1-2-z-STFT-1kHz.png}
        \caption{M1-2}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/time-frequency-spectrum/P1-3-z-STFT-1kHz.png}
        \caption{P1-3}
    \end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/time-frequency-spectrum/P3-3-z-STFT-1kHz.png}
        \caption{P3-3}
    \end{subfigure}
    \caption{Time frequency spectrum}
\end{figure}

\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/time-frequency-spectrum/P1-slow-down.png}
        \caption{Pump P1 turns off}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.48\textwidth}
        \includegraphics[width=\textwidth]{assets/results/time-frequency-spectrum/P2-speed-up.png}
        \caption{Pump P2 turn on}
    \end{subfigure}
\end{figure}

\subsection{Bearing defect frequencies}
% Domain expert analysis
% Bearing frequencies table
\begin{table}[h]
\centering
\begin{tabular}{|l|r|r|r|}
\hline
\textbf{Placement}     & \multicolumn{1}{l|}{\textbf{Motor \#1}} & \multicolumn{1}{l|}{\textbf{Motor \#2}} & \multicolumn{1}{l|}{\textbf{Pump \#3 \& \#4}} \\ \hline
\textbf{Bearing}       & \multicolumn{1}{l|}{6319-C3}            & \multicolumn{1}{l|}{6324-C3}            & \multicolumn{1}{l|}{6317-2Z}                  \\ \hline
\textbf{$n$}           & 8                                       & 8                                       & 8                                             \\ \hline
\textbf{$f_s$}         & 1493                                    & 1493                                    & 1493                                          \\ \hline
\textbf{$d$ {[}mm{]}}  & 33.12                                   & 41.28                                   & 30.00                                         \\ \hline
\textbf{$D$ {[}mm{]}}  & 147.5                                   & 190.0                                   & 132.5                                         \\ \hline
\textbf{$\beta$}       & 0                                       & 0                                       & 0                                             \\ \hline
\textbf{RPM {[}Hz{]}}  & 24.88                                   & 24.88                                   & 24.88                                         \\ \hline
\textbf{BPFO {[}Hz{]}} & 77.18                                   & 77.91                                   & 77.00                                         \\ \hline
\textbf{BPFI {[}Hz{]}} & 121.88                                  & 121.16                                  & 122.07                                        \\ \hline
\textbf{BSF {[}Hz{]}}  & 58.20                                   & 59.97                                   & 57.77                                         \\ \hline
\textbf{FTF {[}Hz{]}}  & 9.65                                    & 9.74                                    & 9.63                                          \\ \hline
\end{tabular}
\end{table}

% Bearing defect frequencies
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/motors.png}
        \caption{Motors}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/motors-dB.png}
        \caption{Pumps}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/pumps.png}
        \caption{Motors}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.24\textwidth}
        \includegraphics[width=\textwidth]{assets/results/defects/pumps-dB.png}
        \caption{Pumps dB}
    \end{subfigure}
\end{figure}

\subsection{Pump long-term monitoring}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{assets/design/sensor/ksb-cloud.jpg}
    \caption{KSB device}
\end{figure}

% Total vibration levels - challange - long-live span (5 years)
\begin{figure}[h]
    \centering
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/results/ksb-cloud/p1.png}
        \caption{Pump P1}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.49\textwidth}
        \includegraphics[width=\textwidth]{assets/results/ksb-cloud/p2.png}
        \caption{Pump P2}
    \end{subfigure}
    \caption{Vibration levels}
\end{figure}

% Reverse engineer of spectra (low resolution)
\begin{figure}[h]
    \centering
    \includegraphics[width=0.8\textwidth]{assets/results/ksb-cloud/spectrum.png}
    \caption{Frequency spectrum}
\end{figure}