### 2024-04-30 thesis/chapters/04-implementation.tex

\chapter{Implementation} \label{chapter:implementation}

\section{Data analysis}
% Jupyter notebook - TSFEL, numpy, pandas (data handling),  scipy (signal processing and stats),   | sklearn, imblearn | matplotlib

\section{Firmware}
% In C, ESP-IDF SDK and FreeRTOS tasks
% Found driver online
% Problems with timing - binary write (osciloscope pictures)
% tool for convert bin2csv

\begin{figure}[h]
    \centering
    \includegraphics[width=0.5\textwidth]{assets/design/sensor/data-logger.jpg}
    \caption{Accelerometer Data Logger}
\end{figure}