### 2024-05-02 thesis/chapters/appendix/B-technical-documentation.tex

\thispagestyle{empty}
\chapter{Technical documentation} \label{appendix:technical-docs}
\pagenumbering{arabic}
\renewcommand*{\thepage}{B-\arabic{page}}

% Jupyter notebooks

% Python function in package (Sphinx)

% C function of firmware