### 2024-05-11 DP_MiroslavHajek-zip/docs/firmware/latex/group__button.tex
Export documentation from Sphinx and Doxygen

\doxysubsubsection{Button}
\hypertarget{group__button}{}\label{group__button}\index{Button@{Button}}

Configuation of the recording button.

\doxysubsubsubsubsection*{Functions}
\begin{DoxyCompactItemize}
\item 
void \mbox{\hyperlink{group__button_ga6a537b9fc1207aed1c9c0a96b325f8c6}{switch\+\_\+enable}} (bool on, gpio\+\_\+isr\+\_\+t isr\+\_\+handler)
\begin{DoxyCompactList}\small\item\em Configure GPIO input pin and interrupt handler for button press. \end{DoxyCompactList}\item 
\Hypertarget{group__button_ga46afed7fc6e15a5c6d9475e46997b3da}\label{group__button_ga46afed7fc6e15a5c6d9475e46997b3da} 
void {\bfseries switch\+\_\+disable} (void)
\begin{DoxyCompactList}\small\item\em Remove interrupt handler for button press. \end{DoxyCompactList}\end{DoxyCompactItemize}

\doxysubsubsubsection{Detailed Description}
Configuation of the recording button.

\doxysubsubsubsection{Function Documentation}
\Hypertarget{group__button_ga6a537b9fc1207aed1c9c0a96b325f8c6}\label{group__button_ga6a537b9fc1207aed1c9c0a96b325f8c6} 
\index{Button@{Button}!switch\_enable@{switch\_enable}}
\index{switch\_enable@{switch\_enable}!Button@{Button}}
\doxyparagraph{\texorpdfstring{switch\_enable()}{switch\_enable()}}
{\footnotesize\ttfamily void switch\+\_\+enable (\begin{DoxyParamCaption}\item[{bool}]{on,  }\item[{gpio\+\_\+isr\+\_\+t}]{isr\+\_\+handler }\end{DoxyParamCaption})}

Configure GPIO input pin and interrupt handler for button press.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em on} & decides whether the button is enabled or disabled \\
\hline
\mbox{\texttt{ in}}  & {\em isr\+\_\+handler} & handler function for button press in interrupt context \\
\hline
\end{DoxyParams}