### 2024-05-11 DP_MiroslavHajek-zip/docs/firmware/latex/group__led.tex

\doxysubsubsection{LED}
\hypertarget{group__led}{}\label{group__led}\index{LED@{LED}}

Configuation of the indicator LED light.

\doxysubsubsubsubsection*{Functions}
\begin{DoxyCompactItemize}
\item 
\Hypertarget{group__led_gaf7f119ca34573b73818f4acfed6d97f6}\label{group__led_gaf7f119ca34573b73818f4acfed6d97f6} 
void {\bfseries led\+\_\+enable} (void)
\begin{DoxyCompactList}\small\item\em Configure GPIO for LED to output mode. \end{DoxyCompactList}\item 
void \mbox{\hyperlink{group__led_gaf3828ebdb3bbb5daebe93e79689fa36b}{led\+\_\+light}} (bool on)
\begin{DoxyCompactList}\small\item\em Set the LED state. \end{DoxyCompactList}\end{DoxyCompactItemize}

\doxysubsubsubsection{Detailed Description}
Configuation of the indicator LED light.

\doxysubsubsubsection{Function Documentation}
\Hypertarget{group__led_gaf3828ebdb3bbb5daebe93e79689fa36b}\label{group__led_gaf3828ebdb3bbb5daebe93e79689fa36b} 
\index{LED@{LED}!led\_light@{led\_light}}
\index{led\_light@{led\_light}!LED@{LED}}
\doxyparagraph{\texorpdfstring{led\_light()}{led\_light()}}
{\footnotesize\ttfamily void led\+\_\+light (\begin{DoxyParamCaption}\item[{bool}]{on }\end{DoxyParamCaption})}

Set the LED state.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em on} & turns LED light to be either on or off \\
\hline
\end{DoxyParams}