### 2024-05-11 DP_MiroslavHajek-zip/docs/firmware/latex/group__storage.tex

\doxysubsubsection{Memory card}
\hypertarget{group__storage}{}\label{group__storage}\index{Memory card@{Memory card}}

Filesystem operations of the SD card.

\doxysubsubsubsubsection*{Functions}
\begin{DoxyCompactItemize}
\item 
sdmmc\+\_\+card\+\_\+t \texorpdfstring{$\ast$}{*} \mbox{\hyperlink{group__storage_ga79585e9fc7f44b32f4e1a653de5bbe3d}{storage\+\_\+enable}} (const char \texorpdfstring{$\ast$}{*}mount\+\_\+point)
\begin{DoxyCompactList}\small\item\em Configure SD/\+MMC bus and mount SD memory card to FAT filesystem. \end{DoxyCompactList}\item 
void \mbox{\hyperlink{group__storage_ga29baf561c143b595123045253ab58f45}{storage\+\_\+disable}} (sdmmc\+\_\+card\+\_\+t \texorpdfstring{$\ast$}{*}\mbox{\hyperlink{group__main_gad85cb86f6cbc94b88b09c3315c9028f1}{card}}, const char \texorpdfstring{$\ast$}{*}mount\+\_\+point)
\begin{DoxyCompactList}\small\item\em Disable and unmount SD memory card from FAT filesystem. \end{DoxyCompactList}\item 
void \mbox{\hyperlink{group__storage_gae29b193e928219e1f41be6b1ac1abea5}{get\+\_\+recording\+\_\+filename}} (char \texorpdfstring{$\ast$}{*}filename, const char \texorpdfstring{$\ast$}{*}path)
\begin{DoxyCompactList}\small\item\em Get file name for new recording with sequentially higher unused number. \end{DoxyCompactList}\end{DoxyCompactItemize}

\doxysubsubsubsection{Detailed Description}
Filesystem operations of the SD card.

\doxysubsubsubsection{Function Documentation}
\Hypertarget{group__storage_gae29b193e928219e1f41be6b1ac1abea5}\label{group__storage_gae29b193e928219e1f41be6b1ac1abea5} 
\index{Memory card@{Memory card}!get\_recording\_filename@{get\_recording\_filename}}
\index{get\_recording\_filename@{get\_recording\_filename}!Memory card@{Memory card}}
\doxyparagraph{\texorpdfstring{get\_recording\_filename()}{get\_recording\_filename()}}
{\footnotesize\ttfamily void get\+\_\+recording\+\_\+filename (\begin{DoxyParamCaption}\item[{char \texorpdfstring{$\ast$}{*}}]{filename,  }\item[{const char \texorpdfstring{$\ast$}{*}}]{path }\end{DoxyParamCaption})}

Get file name for new recording with sequentially higher unused number.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ out}}  & {\em filename} & Available file name for new file \\
\hline
\mbox{\texttt{ in}}  & {\em path} & Base path prefix for saving the recording \\
\hline
\end{DoxyParams}
\Hypertarget{group__storage_ga29baf561c143b595123045253ab58f45}\label{group__storage_ga29baf561c143b595123045253ab58f45} 
\index{Memory card@{Memory card}!storage\_disable@{storage\_disable}}
\index{storage\_disable@{storage\_disable}!Memory card@{Memory card}}
\doxyparagraph{\texorpdfstring{storage\_disable()}{storage\_disable()}}
{\footnotesize\ttfamily void storage\+\_\+disable (\begin{DoxyParamCaption}\item[{sdmmc\+\_\+card\+\_\+t \texorpdfstring{$\ast$}{*}}]{card,  }\item[{const char \texorpdfstring{$\ast$}{*}}]{mount\+\_\+point }\end{DoxyParamCaption})}

Disable and unmount SD memory card from FAT filesystem.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em card} & SD/\+MMC card information structure \\
\hline
\mbox{\texttt{ in}}  & {\em mount\+\_\+point} & path where partition is registered\\
\hline
\end{DoxyParams}
\begin{DoxyReturn}{Returns}
c 
\end{DoxyReturn}
\Hypertarget{group__storage_ga79585e9fc7f44b32f4e1a653de5bbe3d}\label{group__storage_ga79585e9fc7f44b32f4e1a653de5bbe3d} 
\index{Memory card@{Memory card}!storage\_enable@{storage\_enable}}
\index{storage\_enable@{storage\_enable}!Memory card@{Memory card}}
\doxyparagraph{\texorpdfstring{storage\_enable()}{storage\_enable()}}
{\footnotesize\ttfamily sdmmc\+\_\+card\+\_\+t \texorpdfstring{$\ast$}{*} storage\+\_\+enable (\begin{DoxyParamCaption}\item[{const char \texorpdfstring{$\ast$}{*}}]{mount\+\_\+point }\end{DoxyParamCaption})}

Configure SD/\+MMC bus and mount SD memory card to FAT filesystem.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em mount\+\_\+point} & path where the partition will be registered ~\newline
\\
\hline
\end{DoxyParams}
\begin{DoxyReturn}{Returns}
SD/\+MMC card information structure 
\end{DoxyReturn}