### 2024-05-11 DP_MiroslavHajek-zip/docs/firmware/latex/pinout_8h_source.tex

\doxysubsection{pinout.\+h}
\hypertarget{pinout_8h_source}{}\label{pinout_8h_source}\index{firmware/main/include/pinout.h@{firmware/main/include/pinout.h}}

\begin{DoxyCode}{0}
\DoxyCodeLine{00001\ \textcolor{preprocessor}{\#ifndef\ PINOUT\_H}}
\DoxyCodeLine{00002\ \textcolor{preprocessor}{\#define\ PINOUT\_H}}
\DoxyCodeLine{00003\ }
\DoxyCodeLine{00010\ \textcolor{preprocessor}{\#include\ <stdio.h>}}
\DoxyCodeLine{00011\ \textcolor{preprocessor}{\#include\ <string.h>}}
\DoxyCodeLine{00012\ \textcolor{preprocessor}{\#include\ <stdbool.h>}}
\DoxyCodeLine{00013\ }
\DoxyCodeLine{00014\ \textcolor{preprocessor}{\#include\ "{}iis3dwb.h"{}}}
\DoxyCodeLine{00015\ \textcolor{preprocessor}{\#include\ "{}hal/gpio\_types.h"{}}}
\DoxyCodeLine{00016\ \textcolor{preprocessor}{\#include\ "{}driver/sdmmc\_host.h"{}}}
\DoxyCodeLine{00017\ \textcolor{preprocessor}{\#include\ "{}driver/spi\_master.h"{}}}
\DoxyCodeLine{00018\ \textcolor{preprocessor}{\#include\ "{}esp\_err.h"{}}}
\DoxyCodeLine{00019\ \textcolor{preprocessor}{\#include\ "{}esp\_log.h"{}}}
\DoxyCodeLine{00020\ \textcolor{preprocessor}{\#include\ "{}esp\_timer.h"{}}}
\DoxyCodeLine{00021\ \textcolor{preprocessor}{\#include\ "{}freertos/task.h"{}}}
\DoxyCodeLine{00022\ \textcolor{preprocessor}{\#include\ "{}freertos/semphr.h"{}}}
\DoxyCodeLine{00023\ \textcolor{preprocessor}{\#include\ "{}freertos/FreeRTOS.h"{}}}
\DoxyCodeLine{00024\ }
\DoxyCodeLine{00028\ \textcolor{preprocessor}{\#define\ MAX\_FILENAME\ \ \ \ \ \ \ \ \ \ \ \ 256}}
\DoxyCodeLine{00032\ \textcolor{preprocessor}{\#define\ MOUNT\_POINT\ \ \ \ \ \ \ \ \ \ \ \ \ "{}/sd"{}}}
\DoxyCodeLine{00036\ \textcolor{preprocessor}{\#define\ LOG\_FOLDER\ \ \ \ \ \ \ \ \ \ \ \ \ \ MOUNT\_POINT"{}/"{}}}
\DoxyCodeLine{00040\ \textcolor{preprocessor}{\#define\ NO\_WAIT\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 10\ /\ portTICK\_PERIOD\_MS}}
\DoxyCodeLine{00044\ \textcolor{preprocessor}{\#define\ SWITCH\_DEBOUNCE\ \ \ \ \ \ \ \ \ 2000\ /\ portTICK\_PERIOD\_MS}}
\DoxyCodeLine{00045\ }
\DoxyCodeLine{00049\ \textcolor{preprocessor}{\#define\ CARD\_CLK\_PIN\ \ \ \ \ \ \ \ \ \ \ \ 14}}
\DoxyCodeLine{00053\ \textcolor{preprocessor}{\#define\ CARD\_CMD\_PIN\ \ \ \ \ \ \ \ \ \ \ \ 15}}
\DoxyCodeLine{00057\ \textcolor{preprocessor}{\#define\ CARD\_D0\_PIN\ \ \ \ \ \ \ \ \ \ \ \ \ 2}}
\DoxyCodeLine{00058\ }
\DoxyCodeLine{00062\ \textcolor{preprocessor}{\#define\ RECORD\_SWITCH\_PIN\ \ \ \ \ \ \ 34}}
\DoxyCodeLine{00066\ \textcolor{preprocessor}{\#define\ RECORD\_LED\_PIN\ \ \ \ \ \ \ \ \ \ 32}}
\DoxyCodeLine{00067\ }
\DoxyCodeLine{00071\ \textcolor{preprocessor}{\#define\ SENSOR\_MISO\ \ \ \ \ \ \ \ \ \ \ \ \ 13}}
\DoxyCodeLine{00075\ \textcolor{preprocessor}{\#define\ SENSOR\_MOSI\ \ \ \ \ \ \ \ \ \ \ \ \ 16}}
\DoxyCodeLine{00079\ \textcolor{preprocessor}{\#define\ SENSOR\_CLK\ \ \ \ \ \ \ \ \ \ \ \ \ \ 4}}
\DoxyCodeLine{00083\ \textcolor{preprocessor}{\#define\ SENSOR\_CS\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ 5}}
\DoxyCodeLine{00087\ \textcolor{preprocessor}{\#define\ SENSOR\_INT1\ \ \ \ \ \ \ \ \ \ \ \ \ 33}}
\DoxyCodeLine{00091\ \textcolor{preprocessor}{\#define\ SPI\_BUS\_FREQUENCY\ \ \ \ \ \ \ SPI\_MASTER\_FREQ\_8M}}
\DoxyCodeLine{00095\ \textcolor{preprocessor}{\#define\ FIFO\_LENGTH\ \ \ \ \ \ \ \ \ \ \ \ \ 512}}
\DoxyCodeLine{00099\ \textcolor{preprocessor}{\#define\ FIFO\_WATERMARK\ \ \ \ \ \ \ \ \ \ FIFO\_LENGTH\ /\ 2}}
\DoxyCodeLine{00100\ }
\DoxyCodeLine{00104\ \textcolor{preprocessor}{\#define\ NUM\_OF\_FIELDS\ \ \ \ \ \ \ \ \ \ \ 4}}
\DoxyCodeLine{00108\ \textcolor{preprocessor}{\#define\ SENSOR\_SPI\_LENGTH\ \ \ \ \ \ \ NUM\_OF\_FIELDS\ *\ FIFO\_LENGTH}}
\DoxyCodeLine{00112\ \textcolor{preprocessor}{\#define\ QUEUE\_LENGTH\ \ \ \ \ \ \ \ \ \ \ \ 16}}
\DoxyCodeLine{00113\ }
\DoxyCodeLine{00114\ }
\DoxyCodeLine{00124\ \textcolor{preprocessor}{\#define\ SAMPLE\_RATE\ \ \ \ \ \ \ \ \ \ \ \ \ 9000}}
\DoxyCodeLine{00128\ \textcolor{preprocessor}{\#define\ SPI\_BUS\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ \ SPI3\_HOST}}
\DoxyCodeLine{00132\ \textcolor{preprocessor}{\#define\ AUTO\_TURN\_OFF\_US\ \ \ \ \ \ \ \ 60000000}}
\DoxyCodeLine{00136\ \textcolor{preprocessor}{\#define\ ACC\_RESOLUTION\ \ \ \ \ \ \ \ \ \ IIS3DWB\_4g}}
\DoxyCodeLine{00137\ }
\DoxyCodeLine{00141\ \textcolor{keyword}{typedef}\ \textcolor{keyword}{struct\ }\{}
\DoxyCodeLine{00142\ \ \ \ \ uint16\_t\ \mbox{\hyperlink{structAcceleration_a08d970081c96582d6484a1dbb795be88}{len}};\ \ \ \ \ \ \ \ \ \ \ \ \ \ \ }
\DoxyCodeLine{00143\ \ \ \ \ int32\_t\ t[\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\_LENGTH}}];\ \ \ \ \ }
\DoxyCodeLine{00144\ \ \ \ \ int32\_t\ x[\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\_LENGTH}}];\ \ \ \ \ }
\DoxyCodeLine{00145\ \ \ \ \ int32\_t\ y[\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\_LENGTH}}];\ \ \ \ \ }
\DoxyCodeLine{00146\ \ \ \ \ int32\_t\ z[\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\_LENGTH}}];\ \ \ \ \ }
\DoxyCodeLine{00147\ \}\ \mbox{\hyperlink{structAcceleration}{Acceleration}};}
\DoxyCodeLine{00148\ }
\DoxyCodeLine{00157\ \textcolor{keywordtype}{int}\ \mbox{\hyperlink{group__sensor_ga844ea677736e4310ff9a9bee3935330a}{sensor\_enable}}(spi\_device\_handle\_t\ *spi\_dev,\ stmdev\_ctx\_t\ *dev);}
\DoxyCodeLine{00158\ }
\DoxyCodeLine{00164\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__sensor_ga774ff5ef12bc72ef4d96e2ebef38d279}{sensor\_disable}}(spi\_device\_handle\_t\ spi\_dev);}
\DoxyCodeLine{00165\ }
\DoxyCodeLine{00171\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__sensor_ga3a4369193c5acef70192d32950bd4db5}{sensor\_events\_enable}}(stmdev\_ctx\_t\ *dev);}
\DoxyCodeLine{00172\ }
\DoxyCodeLine{00178\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__sensor_gafee0ba0827fd9f252c480faccdee0a97}{sensor\_events\_disable}}(stmdev\_ctx\_t\ *dev);}
\DoxyCodeLine{00179\ }
\DoxyCodeLine{00189\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__datalogger_gae496e529dd495b07dfadcb591230220d}{panic}}(\textcolor{keywordtype}{int}\ delay);}
\DoxyCodeLine{00190\ }
\DoxyCodeLine{00191\ }
\DoxyCodeLine{00205\ sdmmc\_card\_t\ *\mbox{\hyperlink{group__storage_ga79585e9fc7f44b32f4e1a653de5bbe3d}{storage\_enable}}(\textcolor{keyword}{const}\ \textcolor{keywordtype}{char}\ *mount\_point);}
\DoxyCodeLine{00206\ }
\DoxyCodeLine{00215\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__storage_ga29baf561c143b595123045253ab58f45}{storage\_disable}}(sdmmc\_card\_t\ *\mbox{\hyperlink{group__main_gad85cb86f6cbc94b88b09c3315c9028f1}{card}},\ \textcolor{keyword}{const}\ \textcolor{keywordtype}{char}\ *mount\_point);}
\DoxyCodeLine{00216\ }
\DoxyCodeLine{00223\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__storage_gae29b193e928219e1f41be6b1ac1abea5}{get\_recording\_filename}}(\textcolor{keywordtype}{char}\ *filename,\ \textcolor{keyword}{const}\ \textcolor{keywordtype}{char}\ *path);}
\DoxyCodeLine{00224\ }
\DoxyCodeLine{00240\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__button_ga6a537b9fc1207aed1c9c0a96b325f8c6}{switch\_enable}}(\textcolor{keywordtype}{bool}\ on,\ gpio\_isr\_t\ isr\_handler);}
\DoxyCodeLine{00241\ }
\DoxyCodeLine{00245\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__button_ga46afed7fc6e15a5c6d9475e46997b3da}{switch\_disable}}(\textcolor{keywordtype}{void});}
\DoxyCodeLine{00246\ }
\DoxyCodeLine{00259\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__led_gaf7f119ca34573b73818f4acfed6d97f6}{led\_enable}}(\textcolor{keywordtype}{void});}
\DoxyCodeLine{00265\ \textcolor{keywordtype}{void}\ \mbox{\hyperlink{group__led_gaf3828ebdb3bbb5daebe93e79689fa36b}{led\_light}}(\textcolor{keywordtype}{bool}\ on);}
\DoxyCodeLine{00266\ }
\DoxyCodeLine{00272\ \textcolor{preprocessor}{\#endif}}

\end{DoxyCode}