### 2024-05-13 DP_MiroslavHajek/docs/firmware/latex/group__sensor.tex
Thesis send to feedback

\doxysubsubsection{Accelerometer}
\hypertarget{group__sensor}{}\label{group__sensor}\index{Accelerometer@{Accelerometer}}

Configuration of the accelerometer sensor.

\doxysubsubsubsubsection*{Data Structures}
\begin{DoxyCompactItemize}
\item 
struct \mbox{\hyperlink{structAcceleration}{Acceleration}}
\begin{DoxyCompactList}\small\item\em Unprocessed data packet for storing samples from accelerometer. \end{DoxyCompactList}\end{DoxyCompactItemize}
\doxysubsubsubsubsection*{Macros}
\begin{DoxyCompactItemize}
\item 
\Hypertarget{group__sensor_ga4b76a0c2859cfd819a343a780070ee2b}\label{group__sensor_ga4b76a0c2859cfd819a343a780070ee2b} 
\#define {\bfseries SAMPLE\+\_\+\+RATE}~9000
\begin{DoxyCompactList}\small\item\em Interval of the periodic timer in milliseconds that reads out circa half of accelerometer FIFO. \end{DoxyCompactList}\item 
\Hypertarget{group__sensor_gabc52a50e07f79a25a6de181238b9d9aa}\label{group__sensor_gabc52a50e07f79a25a6de181238b9d9aa} 
\#define {\bfseries SPI\+\_\+\+BUS}~SPI3\+\_\+\+HOST
\begin{DoxyCompactList}\small\item\em Hardware bus for accelerometer SPI interface. \end{DoxyCompactList}\item 
\Hypertarget{group__sensor_ga0d440d9222595a142f2d3a4e9e69b11f}\label{group__sensor_ga0d440d9222595a142f2d3a4e9e69b11f} 
\#define {\bfseries AUTO\+\_\+\+TURN\+\_\+\+OFF\+\_\+\+US}~60000000
\begin{DoxyCompactList}\small\item\em Duration of recoding in microseconds (60 s) \end{DoxyCompactList}\item 
\Hypertarget{group__sensor_gac637c5a9d90797796a32b29b70e3a691}\label{group__sensor_gac637c5a9d90797796a32b29b70e3a691} 
\#define {\bfseries ACC\+\_\+\+RESOLUTION}~IIS3\+DWB\+\_\+4g
\begin{DoxyCompactList}\small\item\em Resolution of the accelerometer in a unit of g. \end{DoxyCompactList}\end{DoxyCompactItemize}
\doxysubsubsubsubsection*{Functions}
\begin{DoxyCompactItemize}
\item 
int \mbox{\hyperlink{group__sensor_ga844ea677736e4310ff9a9bee3935330a}{sensor\+\_\+enable}} (spi\+\_\+device\+\_\+handle\+\_\+t \texorpdfstring{$\ast$}{*}spi\+\_\+dev, stmdev\+\_\+ctx\+\_\+t \texorpdfstring{$\ast$}{*}dev)
\begin{DoxyCompactList}\small\item\em Configure SPI bus and set accelerometer to required parameters. \end{DoxyCompactList}\item 
void \mbox{\hyperlink{group__sensor_ga774ff5ef12bc72ef4d96e2ebef38d279}{sensor\+\_\+disable}} (spi\+\_\+device\+\_\+handle\+\_\+t spi\+\_\+dev)
\begin{DoxyCompactList}\small\item\em Remove accelerometer from the SPI bus and disable it. \end{DoxyCompactList}\item 
void \mbox{\hyperlink{group__sensor_ga3a4369193c5acef70192d32950bd4db5}{sensor\+\_\+events\+\_\+enable}} (stmdev\+\_\+ctx\+\_\+t \texorpdfstring{$\ast$}{*}dev)
\begin{DoxyCompactList}\small\item\em Enable accelerometer interrupts. \end{DoxyCompactList}\item 
void \mbox{\hyperlink{group__sensor_gafee0ba0827fd9f252c480faccdee0a97}{sensor\+\_\+events\+\_\+disable}} (stmdev\+\_\+ctx\+\_\+t \texorpdfstring{$\ast$}{*}dev)
\begin{DoxyCompactList}\small\item\em Disable accelerometer interrupts. \end{DoxyCompactList}\end{DoxyCompactItemize}

\doxysubsubsubsection{Detailed Description}
Configuration of the accelerometer sensor.

\doxysubsubsubsection{Function Documentation}
\Hypertarget{group__sensor_ga774ff5ef12bc72ef4d96e2ebef38d279}\label{group__sensor_ga774ff5ef12bc72ef4d96e2ebef38d279} 
\index{Accelerometer@{Accelerometer}!sensor\_disable@{sensor\_disable}}
\index{sensor\_disable@{sensor\_disable}!Accelerometer@{Accelerometer}}
\doxyparagraph{\texorpdfstring{sensor\_disable()}{sensor\_disable()}}
{\footnotesize\ttfamily void sensor\+\_\+disable (\begin{DoxyParamCaption}\item[{spi\+\_\+device\+\_\+handle\+\_\+t}]{spi\+\_\+dev }\end{DoxyParamCaption})}

Remove accelerometer from the SPI bus and disable it.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em spi\+\_\+dev} & SPI bus \\
\hline
\end{DoxyParams}
\Hypertarget{group__sensor_ga844ea677736e4310ff9a9bee3935330a}\label{group__sensor_ga844ea677736e4310ff9a9bee3935330a} 
\index{Accelerometer@{Accelerometer}!sensor\_enable@{sensor\_enable}}
\index{sensor\_enable@{sensor\_enable}!Accelerometer@{Accelerometer}}
\doxyparagraph{\texorpdfstring{sensor\_enable()}{sensor\_enable()}}
{\footnotesize\ttfamily int sensor\+\_\+enable (\begin{DoxyParamCaption}\item[{spi\+\_\+device\+\_\+handle\+\_\+t \texorpdfstring{$\ast$}{*}}]{spi\+\_\+dev,  }\item[{stmdev\+\_\+ctx\+\_\+t \texorpdfstring{$\ast$}{*}}]{dev }\end{DoxyParamCaption})}

Configure SPI bus and set accelerometer to required parameters.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em spi\+\_\+dev} & SPI bus \\
\hline
\mbox{\texttt{ in}}  & {\em dev} & Accelerometer sensor\\
\hline
\end{DoxyParams}
\begin{DoxyReturn}{Returns}
Status code of successful setup 
\end{DoxyReturn}
\Hypertarget{group__sensor_gafee0ba0827fd9f252c480faccdee0a97}\label{group__sensor_gafee0ba0827fd9f252c480faccdee0a97} 
\index{Accelerometer@{Accelerometer}!sensor\_events\_disable@{sensor\_events\_disable}}
\index{sensor\_events\_disable@{sensor\_events\_disable}!Accelerometer@{Accelerometer}}
\doxyparagraph{\texorpdfstring{sensor\_events\_disable()}{sensor\_events\_disable()}}
{\footnotesize\ttfamily void sensor\+\_\+events\+\_\+disable (\begin{DoxyParamCaption}\item[{stmdev\+\_\+ctx\+\_\+t \texorpdfstring{$\ast$}{*}}]{dev }\end{DoxyParamCaption})}

Disable accelerometer interrupts.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em dev} & Accelerometer sensor \\
\hline
\end{DoxyParams}
\Hypertarget{group__sensor_ga3a4369193c5acef70192d32950bd4db5}\label{group__sensor_ga3a4369193c5acef70192d32950bd4db5} 
\index{Accelerometer@{Accelerometer}!sensor\_events\_enable@{sensor\_events\_enable}}
\index{sensor\_events\_enable@{sensor\_events\_enable}!Accelerometer@{Accelerometer}}
\doxyparagraph{\texorpdfstring{sensor\_events\_enable()}{sensor\_events\_enable()}}
{\footnotesize\ttfamily void sensor\+\_\+events\+\_\+enable (\begin{DoxyParamCaption}\item[{stmdev\+\_\+ctx\+\_\+t \texorpdfstring{$\ast$}{*}}]{dev }\end{DoxyParamCaption})}

Enable accelerometer interrupts.

\begin{DoxyParams}[1]{Parameters}
\mbox{\texttt{ in}}  & {\em dev} & Accelerometer sensor \\
\hline
\end{DoxyParams}