### 2024-05-13 DP_MiroslavHajek/docs/firmware/latex/main_8c.tex

\doxysubsection{firmware/main/src/main.c File Reference}
\hypertarget{main_8c}{}\label{main_8c}\index{firmware/main/src/main.c@{firmware/main/src/main.c}}
{\ttfamily \#include "{}pinout.\+h"{}}\newline
\doxysubsubsubsection*{Functions}
\begin{DoxyCompactItemize}
\item 
void {\bfseries push\+\_\+trigger} (void \texorpdfstring{$\ast$}{*}args)
\begin{DoxyCompactList}\small\item\em Task to start or stop recoding after siganl from button press. \end{DoxyCompactList}\item 
void {\bfseries read\+\_\+accelerometer} (void \texorpdfstring{$\ast$}{*}args)
\begin{DoxyCompactList}\small\item\em Task to read FIFO buffer of the accelerometer and write it to Queue. \end{DoxyCompactList}\item 
void {\bfseries write\+\_\+card} (void \texorpdfstring{$\ast$}{*}args)
\begin{DoxyCompactList}\small\item\em Task to write accelerations vectors from Queue to the memory card. \end{DoxyCompactList}\item 
void {\bfseries app\+\_\+main} (void)
\begin{DoxyCompactList}\small\item\em Entrypoint of firmware to setup hardware peripherals and run tasks. \end{DoxyCompactList}\end{DoxyCompactItemize}
\doxysubsubsubsection*{Variables}
\begin{DoxyCompactItemize}
\item 
Task\+Handle\+\_\+t {\bfseries trigger\+\_\+task}
\begin{DoxyCompactList}\small\item\em Task handler for notification of button press. \end{DoxyCompactList}\item 
Task\+Handle\+\_\+t {\bfseries sampler\+\_\+task}
\begin{DoxyCompactList}\small\item\em Task handler for notification from sampling timer. \end{DoxyCompactList}\item 
Queue\+Handle\+\_\+t {\bfseries samples}
\begin{DoxyCompactList}\small\item\em Queue for sending samples from the sensor read task to the memory card write task. \end{DoxyCompactList}\item 
spi\+\_\+device\+\_\+handle\+\_\+t {\bfseries spi}
\begin{DoxyCompactList}\small\item\em SPI bus handle. \end{DoxyCompactList}\item 
stmdev\+\_\+ctx\+\_\+t {\bfseries sensor}
\begin{DoxyCompactList}\small\item\em Accelerometer sensor device. \end{DoxyCompactList}\item 
sdmmc\+\_\+card\+\_\+t \texorpdfstring{$\ast$}{*} {\bfseries card} = NULL
\begin{DoxyCompactList}\small\item\em SD memory card handle. \end{DoxyCompactList}\item 
FILE \texorpdfstring{$\ast$}{*} {\bfseries file} = NULL
\begin{DoxyCompactList}\small\item\em Currently opened file handle. \end{DoxyCompactList}\item 
Semaphore\+Handle\+\_\+t {\bfseries file\+\_\+mutex}
\begin{DoxyCompactList}\small\item\em Mutex to protect file handle. \end{DoxyCompactList}\item 
bool {\bfseries is\+\_\+recording} = false
\begin{DoxyCompactList}\small\item\em Flag for active recording in progress. \end{DoxyCompactList}\item 
int32\+\_\+t {\bfseries sensor\+\_\+timestamp} = 0
\begin{DoxyCompactList}\small\item\em Last seen accelerometer timestamp. \end{DoxyCompactList}\item 
const esp\+\_\+timer\+\_\+create\+\_\+args\+\_\+t \mbox{\hyperlink{group__main_ga6bc19a77e2c149ca9b7e86d2ad25ef64}{sampler\+\_\+timer\+\_\+conf}}
\item 
esp\+\_\+timer\+\_\+handle\+\_\+t {\bfseries sampler\+\_\+timer}
\begin{DoxyCompactList}\small\item\em Periodic timer to signal when to read FIFO buffer from accelerometer. \end{DoxyCompactList}\item 
const esp\+\_\+timer\+\_\+create\+\_\+args\+\_\+t \mbox{\hyperlink{group__main_gacb077cf1ef581c961e7e5222a98f6897}{stop\+\_\+timer\+\_\+conf}}
\item 
esp\+\_\+timer\+\_\+handle\+\_\+t {\bfseries stop\+\_\+timer}
\begin{DoxyCompactList}\small\item\em Timer to stop recording after fixed amount of time. \end{DoxyCompactList}\end{DoxyCompactItemize}

\doxysubsubsection{Detailed Description}
Application setup