### 2024-05-13 DP_MiroslavHajek/docs/firmware/latex/structAcceleration.tex

\doxysubsection{Acceleration Struct Reference}
\hypertarget{structAcceleration}{}\label{structAcceleration}\index{Acceleration@{Acceleration}}

Unprocessed data packet for storing samples from accelerometer.

{\ttfamily \#include $<$pinout.\+h$>$}

\doxysubsubsubsection*{Data Fields}
\begin{DoxyCompactItemize}
\item 
\Hypertarget{structAcceleration_a08d970081c96582d6484a1dbb795be88}\label{structAcceleration_a08d970081c96582d6484a1dbb795be88} 
uint16\+\_\+t {\bfseries len}
\begin{DoxyCompactList}\small\item\em Number of samples in every array in the structure. \end{DoxyCompactList}\item 
\Hypertarget{structAcceleration_ab7a417b34f3fc19dc3ae14fa5ac4c0b6}\label{structAcceleration_ab7a417b34f3fc19dc3ae14fa5ac4c0b6} 
int32\+\_\+t {\bfseries t} \mbox{[}\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\+\_\+\+LENGTH}}\mbox{]}
\begin{DoxyCompactList}\small\item\em Array of timestamps relative to time when sensor was enabled. \end{DoxyCompactList}\item 
\Hypertarget{structAcceleration_a4744bf8f88deaa5ae9f2e58ebb244dc7}\label{structAcceleration_a4744bf8f88deaa5ae9f2e58ebb244dc7} 
int32\+\_\+t {\bfseries x} \mbox{[}\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\+\_\+\+LENGTH}}\mbox{]}
\begin{DoxyCompactList}\small\item\em Accelerometer samples for X axis. \end{DoxyCompactList}\item 
\Hypertarget{structAcceleration_abe8b39739af95fa91287e19c9b9ac3c6}\label{structAcceleration_abe8b39739af95fa91287e19c9b9ac3c6} 
int32\+\_\+t {\bfseries y} \mbox{[}\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\+\_\+\+LENGTH}}\mbox{]}
\begin{DoxyCompactList}\small\item\em Accelerometer samples for Y axis. \end{DoxyCompactList}\item 
\Hypertarget{structAcceleration_aba0250b0210d700100bccaa29ff4ffa4}\label{structAcceleration_aba0250b0210d700100bccaa29ff4ffa4} 
int32\+\_\+t {\bfseries z} \mbox{[}\mbox{\hyperlink{group__datalogger_gaf759d6c69e76948f9933f3872d997adb}{FIFO\+\_\+\+LENGTH}}\mbox{]}
\begin{DoxyCompactList}\small\item\em Accelerometer samples for Z axis. \end{DoxyCompactList}\end{DoxyCompactItemize}

\doxysubsubsection{Detailed Description}
Unprocessed data packet for storing samples from accelerometer.

The documentation for this struct was generated from the following file\+:\begin{DoxyCompactItemize}
\item 
firmware/main/include/pinout.\+h\end{DoxyCompactItemize}