### 2024-05-13 DP_MiroslavHajek/docs/firmware/latex/topics.tex

\doxysubsection{Topics}
Here is a list of all topics with brief descriptions\+:\begin{DoxyCompactList}
\item \contentsline{section}{Accelerometer Data logger}{\pageref{group__datalogger}}{}
\begin{DoxyCompactList}
\item \contentsline{section}{Accelerometer}{\pageref{group__sensor}}{}
\item \contentsline{section}{Memory card}{\pageref{group__storage}}{}
\item \contentsline{section}{Button}{\pageref{group__button}}{}
\item \contentsline{section}{LED}{\pageref{group__led}}{}
\end{DoxyCompactList}
\item \contentsline{section}{Firmware Tasks}{\pageref{group__main}}{}
\end{DoxyCompactList}